

% Advanced Topics in Communication Networks D-ITET
% ===========================================================================
% @Author: Noah Huetter
% @Date:   2019-09-24 17:26:28
% @Last Modified by:   noah
% @Last Modified time: 2019-10-08 10:59:46
% ---------------------------------------------------------------------------

\documentclass[a4paper, fontsize=8pt, DIV=1]{scrartcl}
\usepackage{lastpage}
\usepackage{hyperref}
% Include general settings and customized commands
\input{settings/general}
\input{settings/commands}

% This package makes formulas a bit more compact but less beautiful
% \usepackage{newtxtext,newtxmath}

% scala language description
\lstdefinelanguage{BNF}{%
    alsoletter={-},%
    sensitive,%
}[keywords,comments]%


% scala language description
\lstdefinelanguage{P4}{%
    sensitive,%
    backgroundcolor=\color{codeblue},%
%    morecomment=[l]//,%
%    morecomment=[s]{/*}{*/},%
}[keywords,comments]%

\definecolor{codeblue}{HTML}{DEF0FE}
\lstdefinestyle{P4style}{
    language=P4,%
    frame=none,%
    backgroundcolor=\color{codeblue},%
    keywords={action, action_function_declaration, action_profile, action_selector, algorithm, and, apply, attribute, attributes, bit, bytes, bytes_and_packets, calculated_field, control, counter, direct, dynamic_action_selection, else, extern, extern_type, extract, false, field_list, field_list_calculation, fields, header, header_type, hit, if, in, inout, input, instance_count, int, last, layout, mask, max, metadata, meter, method, min, min_width, miss, next, not, optional, or, output_width, packets, parse_error, parser, parser_drop, parser_exception, parser_value_set, primitive_action_declaration, range, register, result, return, saturating, select, selection_key, set_metadata, signed, static, table, true, update, valid, varbit, verify, width},%
    basicstyle=\ttfamily,%
    aboveskip=3mm,%
    belowskip=3mm,%
    fontadjust=true,%
    keepspaces=true,%
    keywordstyle=\bfseries,%
    captionpos=b,%
    framerule=0.3pt,%
    firstnumber=0,%
    numbersep=1.5mm,%
    numberstyle=\tiny,%
}
\lstset{%
    basicstyle=\ttfamily,%
%    language=P4,%
    aboveskip=3mm,%
    belowskip=3mm,%
    fontadjust=true,%
%    columns=[c]fixed,%
    keepspaces=true,%
%    commentstyle=\itshape,%
    keywordstyle=\bfseries,%
    captionpos=b,%
    framerule=0.3pt,%
    firstnumber=0,%
    numbersep=1.5mm,%
    numberstyle=\tiny,%
}

% \bibliography{semiconductordevices}
% \bibliographystyle{ieeetr}
\medmuskip=1mu

%change page style for header
\pagestyle{fancy}
\footskip 20pt

% Uncomment this line to make formulasheet ultra compact
% This removes
% - list of variables
% \newcommand{\makeultracompact}{irrelevant}
\let\makeultracompact\undefined

% Make stuff ultra compact if so desired
\ifdefined\makeultracompact
  \setlength{\parskip}{0pt}
  \setlength{\abovedisplayskip}{0pt}
  \setlength{\belowdisplayskip}{0pt}
  \setlength{\abovedisplayshortskip}{0pt}
  \setlength{\belowdisplayshortskip}{0pt}
\else
\fi
 
% -----------------------------------------------------------------------
\IfFileExists{../build/revision.tex}{
  \input{../build/revision.tex}
  \rhead{Compiled: \compiledate \hspace{1em} on: \hostname \hspace{1em} git-sha: \revision \hspace{1em} Noah Huetter}
}{\rhead{Noah Huetter}}

\ifdefined\makeultracompact
  \lhead{ETH Advanced Topics in Communication Networks 2019 \hspace{1em}compact version}
\else
  \lhead{ETH Advanced Topics in Communication Networks 2019}
\fi
\chead{\thepage}
\cfoot{}
\headheight 17pt \headsep 10pt
\title{ETH Advanced Topics in Communication Networks 2019}
\author{Noah Huetter}

\date{\today}
\begin{document}

\setcounter{page}{0}
\setcounter{secnumdepth}{2} %no enumeration of sections
\begin{multicols*}{2}
	\section*{Disclaimer}
	This summary is part of the lecture ``ETH Advanced Topics in Communication Networks'' (227-0575-00L) by Prof. Dr. Laurent Vanbever (FS19). It is based on the lecture. \\[6pt]
	Please report errors to \href{mailto:huettern@student.ethz.ch}{huettern@student.ethz.ch} such that others can benefit as well.\\[6pt]	
  The upstream repository can be found at \href{https://github.com/noah95/formulasheets}{https://github.com/noah95/formulasheets}
	\vfill\null
  \columnbreak
  %%%%%%%%%%%%%%%%%%%%%%%%%%%%%
  \tableofcontents
  \vfill\null
  %\columnbreak
  %%%%%%%%%%%%%%%%%%%%%%%%%%%%%
	\pagebreak
  \maketitle 
  \setcounter{page}{1}
  \thispagestyle{fancy}

  % ---------------------------------------------------------------------------
  \section{P4 language}
  % ---------------------------------------------------------------------------
  \subsection{Data types}

  \subsubsection{Base types}

  \begin{tabularx}{\linewidth}{ l X}
    \lstinline[style=P4style]!bool! &
    Boolean value\\
    \lstinline[style=P4style]!bit<W>! &
    Bit-string of width \texttt{W}\\
    \lstinline[style=P4style]!int<W>! &
    Signed integer of width \texttt{W}\\
    \lstinline[style=P4style]!varbit<W>! &
    Bit-string of dynamiv length \texttt{<=W}\\
    \lstinline[style=P4style]!match_kind! &
    describes ways to match table keys\\
    \lstinline[style=P4style]!error! &
    used to signal errors\\
    \lstinline[style=P4style]!void! &
    no values, used in few restricted circumstances\\
  \end{tabularx}

  \subsubsection{Header}
  Parsing a packet using \texttt{extract()} fills in the fields of the header.
  A successful \texttt{extract()} sets to true the validity bit of the extraced header.

  \begin{lstlisting}[style=P4style]
action:
header Ethernet_h {
  bit<48> dstAddr;
  bit<48> srcAddr;
  bit<16> etherType;
}
Ethernet_h ethernetHeader;
\end{lstlisting}

  \subsubsection{Header Stack}
  Array of up to \texttt{N} headers of type \texttt{Mpls\_h}.

\begin{lstlisting}[style=P4style]
header Mpls_h {
  bit<20> label;
  bit<3>  tc;
  bit     bos;
  bit<8>  ttl;
}
Mpls_h[10] mpls;
\end{lstlisting}

  \subsubsection{Header union}
Either \texttt{IPv4} or \texttt{IPv6} header is present.

\begin{lstlisting}[style=P4style]
header_union IP_h {
 IPv4_h v4;
 IPv6_h v6;
}
\end{lstlisting}


  \subsubsection{Struct}
  Unordered collection of named members.

\begin{lstlisting}[style=P4style]
struct standard_metadata_t {
 bit<9> ingress_port;
 bit<9> egress_spec;
 bit<9> egress_port;
 ...
}
\end{lstlisting}

  \subsubsection{Tuple}
  Unordered collection of unnamed members.

\begin{lstlisting}[style=P4style]
tuple<bit<32>, bool> x;
x = { 10, false }
\end{lstlisting}

  \subsubsection{Other}
  \begin{tabularx}{\linewidth}{ l X}
  \lstinline[style=P4style]!enum Prio \{High, Low\}! &
  Enumeration \\
  \lstinline[style=P4style]!typedef bit<48> macAddr_t! &
  Create subtypes \\
  \lstinline[style=P4style]!extern! &
  {}  \\
  \lstinline[style=P4style]!parser! &
  {}  \\
  \lstinline[style=P4style]!control! &
  {}  \\
  \lstinline[style=P4style]!package! &
  {}  \\
  \end{tabularx}

  \subsection{Data structures}

  \begin{lstlisting}[style=P4style]
struct standard_metadata_t {
  bit<9>  ingress_port;
  bit<9>  egress_spec;
  bit<9>  egress_port;
  bit<32> instance_type;
  bit<32> packet_length;
  bit<32> enq_timestamp;
  bit<19> enq_qdepth;
  bit<32> deq_timedelta;
  bit<19> deq_qdepth;

  // intrinsic to metadata v1model
  bit<48> ingress_global_timestamp;
  bit<48> egress_global_timestamp;
  bit<16> mcast_grp;
  bit<16> egress_rid;
  bit<1>  checksum_error;
  error parser_error;
  bit<3> priority;
}\end{lstlisting}

\begin{tabularx}{\linewidth}{ l X }
  \lstinline[style=P4style]!ingress_port! &
  Port the packet was received \\
  \lstinline[style=P4style]!egress_spec! &
  Port where the packet will be transmitted \\
\end{tabularx}

  \subsection{Examples}

  \subsubsection{Multicast group ID}
  For a L2-switch to work, it has to multicast packets. Simple switch has a multicast features that
  allows setting a metadata field so that packets are sent to multiple ports.

  First, on the switch, create a multicast group, node and assign these.
  \begin{lstlisting}[language=bash]
mc_mgrp_create <mcast_grp_id>
mc_node_create <replication_id> <port_number> [port_number]
mc_node_associate <mcast_grp_id> <node_handle_id>\end{lstlisting}
  \texttt{mc\_node\_create} returns the \texttt{node\_handle\_id}. For every \texttt{replication\_id} the \texttt{mc\_node\_create} is incremented.

  Example: Mulcicast \texttt{mcast\_grp\_id=1} with \texttt{replication\_id=0} on ports \texttt{1,2,3,4} 
  \begin{lstlisting}[language=bash]
mc_mgrp_create 1
mc_node_create 0 1 2 3 4
mc_node_associate 1 0\end{lstlisting}

  To send a packet to this group, in P4 set the \texttt{standard\_metadata.mcast\_grp} to \texttt{mcast\_grp\_id} during the ingress pipeline.

  \subsubsection{Cloning Packets}
  Is a feature of simple switch.

  Configure the switch with a mirroring session \texttt{session} mapped to an \texttt{output\_port}.
  \begin{lstlisting}[style=P4style]
mirroring_add <session> <output_port>\end{lstlisting}

  During ingress pipeline, a packet can be cloned using \texttt{clone}. The packet will be cloned to the egress pipeline and port configured by \texttt{mirroring\_add}. When using \texttt{clone3} you can add as a third parameter a metadata \texttt{struct}.
  \begin{lstlisting}[style=P4style]
clone(CloneType.I2E, <session>)
// Generally
clone(in CloneType type, in bit<32> session)
clone3<T>(in CloneType type, in bit<32> session, in T data)\end{lstlisting}

  To identify a cloned packet in the egress pipeline, check the \texttt{standard\_metadata.instance\_type==1} field.

  When a packet is cloned all its metadata fields
  are reset to the default value (usually 0).


  \subsubsection{Digest packets}
  A feature of simple switch to send packets to the controller without having to clone it.

  \begin{lstlisting}[style=P4style]
// Define a struct that is sent to the controller
struct learn_t {
    macAddr_t srcAddr;
    inPort_t  inPort;
}
// Add the struct to the metadata
struct metadata {
    /* empty */
    //TODO 3: delcare one learn_t variable
    learn_t learn;
}
// fill in the meatdata fields
meta.learn.srcAddr = hdr.ethernet.srcAddr;
meta.learn.inPort = (inPort_t)standard_metadata.ingress_port;
// send to the controller
// nothing else has to be done
digest(1, meta.learn); // first argument is always 1\end{lstlisting}

  
\end{multicols*}

\setcounter{secnumdepth}{2}
\end{document}
